deckard: A Declarative Tool for Machine Learning Robustness Evaluations

### 2025-12-09 paper.tex
Commit from GitHub Actions (Render R Markdown)

The software package presented, called \texttt{deckard}, is a modular
software toolkit designed to streamline and standardize experimentation
in machine learning (ML) with a particular focus on the adversarial
scenario. It provides a flexible, extensible framework for defining,
executing, and analysing end-to-end ML pipelines in the context of a
malicious actor. As it is built on top of the Hydra configuration
system, deckard supports declarative YAML-based configuration of data
preprocessing, model training, and adversarial attack pipelines,
enabling reproducible, framework-agnostic experimentation across diverse
ML settings.

In addition to configuration management, \texttt{deckard} includes a
suite of utilities for distributed and parallel execution, automated
hyperparameter optimisation, visualisation, and result aggregation. The
tooling abstracts away much of the engineering overhead typically
involved in adversarial ML research, allowing researchers to focus on
algorithmic insights rather than implementation details. The presented
software facilitates rigorous benchmarking by maintaining an auditable
trace of configurations, random seeds, and intermediate outputs
throughout the experimental lifecycle.

The system is compatible with a variety of ML frameworks and several
classes of adversarial attacks, making it a suitable backend for both
large-scale automated testing and fine-grained empirical analysis. By
providing a unified interface for experimental control, \texttt{deckard}
accelerates the development and evaluation of robust models, and helps
close the gap between research prototypes and verifiable, reproducible
results.

@@

While tools such as \texttt{mlflow} (Zaharia et al. 2018), Weights \&
Biases (Biewald 2020), \texttt{optuna} (Akiba et al. 2019), and
Kubernetes (Kubernetes 2019) provide essential infrastructure for model
tracking and experiment management, \texttt{deckard} occupies a
different position in the ML ecosystem---focusing specifically on
configurable, adversarially robust experimentation.

Unlike MLflow and Weights \& Biases, which emphasize logging,
visualization, and reproducibility for various ML frameworks, deckard
enforces reproducibility by construction through its declarative,
YAML-driven configuration system built on Facebook's hydra (Yadan 2019)
configuration management tool. In contrast to cloud-management software
like Kubernetes---which is a general-purpose container orchestration
platform---\texttt{deckard} abstracts away orchestration details and
offers native support for parallel and distributed experimentation,
tailored to ML workflows involving attack/defense cycles, model
retraining, or optimisation. While deckard integrates tightly with IBM's
Adversarial Robustness Toolbox (Nicolae et al. 2018), the software is
designed to be easily extensible to other attack frameworks. The human-
and machine-readable parameter configuration system allows researchers
to declaratively define end-to-end pipelines that span data sampling,
preprocessing, model training, attack generation, defense evaluation,
multi-objective optimisation, and visualisation. Tools like \texttt{ray}
(Moritz et al. 2018), \texttt{optuna} (Akiba et al. 2019), or
\texttt{nevergrad} (Bennet et al. 2021) offer components of this
pipeline (\emph{e.g.}, hyperparameter search or configuration
management), but lack unified support for adversarial ML, verification,
or auditability at scale. While \texttt{deckard} complements these
existing tools, and in many cases can be integrated with them, its
primary contribution is in automating and verifying adversarial ML
experiments in a way that is both extensible and framework-agnostic.

@@

Various versions of this software have been used in several recently
published and not-yet-published works by the author of this paper, all
of which are available in the \texttt{examples} folder in the source
code repository \url{https://github.com/simplymathematics/deckard}. One
published work, now reproducible via the
\texttt{examples/attack\_defence\_survey} folder, includes a large
survey of attacks and defences against canonical datasets and models (C.
Meyers, Löfstedt, and Elmroth 2023). Another work analysed the run-time
requirements of attacks against a particular model before and after
retraining against those attacks (Charles Meyers et al. 2024)
(reproducible via \texttt{examples/retraining}). The next paper
formalised a method for estimating the time-to-failure of a given model
against a suite of attacks and introduce a metric that quantifies the
ratio of attack and training cost (Meyers et al. 2023) (reproducible via
\texttt{examples/survival\_heuristic}). Furthemore, a not yet published
work uses this time-to-failure model as a mechanism for analysing the
cost efficacy of various hardware choices in the context of adversarial
attacks (reproducible via \texttt{examples/power}) (C. Meyers et al.
2024). Another work exploits the tooling to train a custom model that is
designed to run client-side by using compression algorithms to measure
the distance between text (reproducible via
\texttt{examples/compression}).

@@

Since ML projects can exploit specialized hardware such as multi-core
processors or GPUs, and often rely on clusters of machines for
large-scale data processing, it was necessary to enable parallel and
distributed experiment execution and model optimization. By leveraging
the \texttt{hydra} configuration framework, \texttt{deckard}
automatically supports optimization libraries like \texttt{nevergrad}
(Bennet et al. 2021), \texttt{Adaptive\ Experimentation} (A. Developers
2025), and \texttt{optuna} (Akiba et al. 2019), making the software
modular and extensible. Additionally, experiments can be managed using a
variety of popular job schedulers, including \texttt{Ray} (Moritz et al.
2018), \texttt{Redis\ Queue} (Stamps 2025), and \texttt{slurm} (Yoo,
Jette, and Grondona 2003) for distributed jobs or \texttt{joblib} (Joblib
Developers 2025) for jobs on a single machine.

By using a declarative design, a given set of experiments can be
specified once and executed seamlessly across different back-ends without
modification to the underlying codebase. This makes \texttt{deckard}
both adaptable and scalable, suitable for use on personal laptops,
multi-node servers, or large-scale, high-performance clusters. When
configured appropriately, experiment batches can be parallelized,
enabling massive parameter sweeps, ensemble evaluations, or adversarial
robustness tests to be executed in parallel---reducing turnaround time
while maintaining strong guarantees on reproducibility and auditability.
The design of the presented software prioritizes clarity and
maintainability by capturing each experimental configuration as a YAML
artifact, making both successful and failed runs equally traceable and
shareable. This approach transforms experiment tracking from an
afterthought into a first-class component of the trustworthy ML
workflow.

@@

\leavevmode\vadjust pre{\hypertarget{ref-wandb}{}}%
Biewald, Lukas. 2020. {``Experiment Tracking with Weights and Biases.''}
\url{https://www.wandb.com/}.

@@

@@

\leavevmode\vadjust pre{\hypertarget{ref-k8s}{}}%
Kubernetes. 2019. {``Kubernetes--an Open Source System for Managing
Containerized Applications.''} Github.
\url{https://github.com/kubernetes/kubernetes}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-art}{}}%
Nicolae, Maria-Irina, Mathieu Sinn, Minh Ngoc Tran, Beat Buesser,
Ambrish Rawat, Martin Wistuba, Valentina Zantedeschi, et al. 2018.
{``Adversarial Robustness Toolbox V1. 0.0.''} \emph{arXiv Preprint
arXiv:1807.01069}.

\leavevmode\vadjust pre{\hypertarget{ref-pytorch}{}}%
Paszke, Adam, Sam Gross, Francisco Massa, Adam Lerer, James Bradbury,
Gregory Chanan, Trevor Killeen, et al. 1912. {``Pytorch: An Imperative
Style, High-Performance Deep Learning Library\@. arXiv 2019.''}
\emph{arXiv Preprint arXiv:1912.01703} 10.

@@

\leavevmode\vadjust pre{\hypertarget{ref-mlflow}{}}%
Zaharia, Matei, Andrew Chen, Aaron Davidson, Ali Ghodsi, Sue Ann Hong,
Andy Konwinski, Siddharth Murching, et al. 2018. {``Accelerating the
Machine Learning Lifecycle with MLflow.''} \emph{IEEE Data Eng. Bull.}
41 (4): 39--45.

>>>>

The software package presented, called \texttt{deckard}, is a modular
toolkit that streamlines and standardizes machine learning (ML)
experimentation, with a particular emphasis on evaluating models under
worst-case perturbations(Goodfellow, Shlens, and Szegedy 2014). It
offers a flexible, extensible framework for defining, running, and
analysing end-to-end ML pipelines in settings where inputs may be
manipulated by an attacker or corrupted by unexpected noise (Biggio et
al. 2013).

Built on the Hydra configuration system (Yadan 2019), \texttt{deckard}
supports declarative YAML-based configuration for data preprocessing,
model training, and adversarial attack pipelines. This makes it easy to
run reproducible, framework-agnostic experiments across a wide range of
ML domains.

Beyond configuration management, \texttt{deckard} provides tools for
distributed and parallel execution, automated hyperparameter
optimization, visualization, and result aggregation. By handling much of
the engineering overhead common in adversarial ML research, it lets
researchers focus on developing ideas rather than juggling
infrastructure. The system also ensures rigorous benchmarking by keeping
an auditable record of configurations, seeds, and intermediate outputs
throughout each experiment.

Compatible with multiple ML frameworks and a variety of attack methods,
\texttt{deckard} works both as a backend for large-scale automated
testing and as a platform for detailed empirical analysis. With its
unified interface for experiment control, the toolkit accelerates the
development of robust models and helps bridge the gap between
exploratory research and reproducible, trustworthy results .

To avoid any ambiguity about contributions, we clarify that
\texttt{deckard} is a method and software system introduced in this
work. Although the implementation is publicly available on GitHub, the
modular design, configuration-driven framework, and experimental
methodology are original to this paper. While the prior works outlined
in the Usage Section can be fully reproduced with this software, this
paper is the first to present the system as a cohesive method, fully
specified, documented, and intended for broad reuse.

@@

The software package presented, called \texttt{deckard}, is a modular
toolkit that streamlines and standardizes machine learning (ML)
experimentation, with a particular emphasis on evaluating models under
worst-case perturbations (Goodfellow, Shlens, and Szegedy 2014). It
offers a flexible, extensible framework for defining, running, and
analysing end-to-end ML pipelines in settings where inputs may be
manipulated by an attacker or corrupted by unexpected noise (Biggio et
al. 2013).

Built on the Hydra configuration system (Yadan 2019), \texttt{deckard}
supports declarative YAML-based configuration for data preprocessing,
model training, and adversarial attack pipelines. This makes it easy to
run reproducible, framework-agnostic experiments across a wide range of
ML domains.

Beyond configuration management, \texttt{deckard} provides tools for
distributed and parallel execution, automated hyperparameter
optimization, visualization, and result aggregation. By handling much of
the engineering overhead common in adversarial ML research, it lets
researchers focus on developing ideas rather than juggling
infrastructure. The system also ensures rigorous benchmarking by keeping
an auditable record of configurations, seeds, and intermediate outputs
throughout each experiment.

Compatible with multiple ML frameworks and a variety of attack methods,
\texttt{deckard} works both as a backend for large-scale automated
testing and as a platform for detailed empirical analysis. With its
unified interface for experiment control, the toolkit accelerates the
development of robust models and helps bridge the gap between
exploratory research and reproducible, trustworthy results .

To avoid any ambiguity about contributions, we clarify that
\texttt{deckard} is a method and software system introduced in this
work. Although the implementation is publicly available on GitHub, the
modular design, configuration-driven framework, and experimental
methodology are original to this paper. While the prior works outlined
in the Usage Section can be fully reproduced with this software, this
paper is the first to present the system as a cohesive method, fully
specified, documented, and intended for broad reuse.

@@

Various versions of this software have been used in several recently
published and not-yet-published works by the author of this paper, all
of which are available in the \texttt{examples} folder in the source
code repository \url{https://github.com/simplymathematics/deckard}. One
published work, now reproducible via the
\texttt{examples/attack\_defence\_survey} folder, includes a large
survey of attacks and defences against canonical datasets and models (C.
Meyers, Löfstedt, and Elmroth 2023). Another work analysed the run-time
requirements of attacks against a particular model before and after
retraining against those attacks (Charles Meyers and Elmroth 2024)
(reproducible via \texttt{examples/retraining}). The next paper
formalised a method for estimating the time-to-failure of a given model
against a suite of attacks and introduce a metric that quantifies the
ratio of attack and training cost (Meyers et al. 2023) (reproducible via
\texttt{examples/survival\_heuristic}). Furthemore, a not yet published
work uses this time-to-failure model as a mechanism for analysing the
cost efficacy of various hardware choices in the context of adversarial
attacks (reproducible via \texttt{examples/power}) (C. Meyers et al.
2024). Another work exploits the tooling to train a custom model that is
designed to run client-side by using compression algorithms to measure
the distance between text (reproducible via
\texttt{examples/compression}).

@@

Since ML projects can exploit specialized hardware such as multi-core
processors or GPUs, and often rely on clusters of machines for
large-scale data processing, it was necessary to enable parallel and
distributed experiment execution and model optimization. By leveraging
the \texttt{hydra} configuration framework, \texttt{deckard}
automatically supports optimization libraries like \texttt{nevergrad}
(Bennet et al. 2021), \texttt{Adaptive\ Experimentation} (A. Developers
2025), and \texttt{optuna} (Akiba et al. 2019), making the software
modular and extensible. Additionally, experiments can be managed using a
variety of popular job schedulers, including \texttt{Ray} (Moritz et al.
2018), \texttt{Redis\ Queue} (Stamps 2025), and \texttt{slurm} (Yoo,
Jette, and Grondona 2003) for distributd jobs or \texttt{joblib} (J.
Developers 2025) for jobs on a single machine.

By using a declarative design, a given set of experiments can be
specified once and executed seamlessly across different backends without
modification to the underlying codebase. This makes \texttt{deckard}
both adaptable and scalable, suitable for use on personal laptops,
multi-node servers, or large-scale, high-performance clusters. When
configured appropriately, experiment batches can be parallelized,
enabling massive parameter sweeps, ensemble evaluations, or adversarial
robustness tests to be executed in parallel---reducing turnaround time
while maintaining strong guarantees on reproducibility and auditability.
The design of the presented software prioritizes clarity and
maintainability by capturing each experimental configuration as a YAML
artifact, making both successful and failed runs equally traceable and
shareable. This approach transforms experiment tracking from an
afterthought into a first-class component of the trustworthy ML
workflow.

@@

\leavevmode\vadjust pre{\hypertarget{ref-biggio_evasion_2013}{}}%
Biggio, Battista, Igino Corona, Davide Maiorca, Blaine Nelson, Nedim
Šrndić, Pavel Laskov, Giorgio Giacinto, and Fabio Roli. 2013. {``Evasion
Attacks Against Machine Learning at Test Time,''} 387--402.

@@

\leavevmode\vadjust pre{\hypertarget{ref-fgm}{}}%
Goodfellow, Ian J, Jonathon Shlens, and Christian Szegedy. 2014.
{``Explaining and Harnessing Adversarial Examples.''}
\emph{arXiv:1412.6572}.

@@

@@

\leavevmode\vadjust pre{\hypertarget{ref-pytorch}{}}%
Paszke, Adam, Sam Gross, Francisco Massa, Adam Lerer, James Bradbury,
Gregory Chanan, Trevor Killeen, et al. 1912. {``Pytorch: An Imperative
Style, High-Performance Deep Learning Library. arXiv 2019.''}
\emph{arXiv Preprint arXiv:1912.01703} 10.

@@
